---
title: "AN588 - Homework 2"
author: "Christian Mei"
editor: visual
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: true         
    toc_depth: 2
---

# Titis in the Mist...

Every Saturday morning, at the same time, a primatologist goes and sits in the forest to listen for titi monkey calls, counting the number of calls they hear in a [2 hour window from 5am to 7am]{.underline}. Based on previous knowledge, she believes that the [mean number of calls she will hear in that time is exactly 15.]{.underline} Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

What is the probability that she will hear more than 8 calls during any given session?

What is the probability that she will hear no calls in a session?

What is the probability that she will hear exactly 3 calls in a session?

Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.

Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).

Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

# Answer Attempt

I decided to answer the questions in the order I think felt more inuitive

## Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30

We first need to create the distribution/dataset that the primatologist collected before plotting it.

```{r}
call_ranges <- 0:30
l = 15 # Poisson mean lambda based on mean number of calls
titis_calls <- dpois(x = call_ranges, lambda = l)


barplot(titis_calls, names.arg = call_ranges, space = 0, xlab = "Number of Calls per Day", ylab = "Pr(X = x)", main = "Probability Mass Function - Titis Calls")
```

## Simulate 104 results from this distribution:

My initial attempt which resulted in a mistake. Why do you think this is the case? Why can't I sample from the original `titis-calls` collection?

```{r}
sim <- NULL  # sets up a dummy variable to hold our 10000 simulations
n <- 104
for (i in 1:length(titis_calls)) {
    sim[i] <- mean(sample(titis_calls, n, replace = TRUE))
}

barplot(sim, names.arg = call_ranges, space = 0, xlab = "Number of Calls per Day", ylab = "Pr(X = x)", main = "Probability Mass Function - Titis Calls")


```

The problem above is that I am by sampling from `titis_calls` , I was sampling probability values and not data points (calls per day). Hence, this method is incorrect

What the assignment was actually asking for was to sample from a Poisson distribution with similar values, e.g same lambda mean = 15.

> BP: I appreciate that you remarked on your initial attempt and what went wrong with it. My first instinct on seeing that prompt was to try and run a for loop too, though I quickly realized like you did that just using the rpois function would be much more efficient.

```{r}
set.seed(1) # So that everytime I run rpois, I get the same set of random numbers
simulated_calls <- rpois(104, lambda = l)
simulated_calls
```

> BP: I also like that you set a seed to ensure consistency of your simulated values if the code chunk is rerun later on, that's something I'll try to work into my approach as well

## Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30.

```{r}
hist(simulated_calls, breaks = 0:30, freq = TRUE, col = "pink",
     xlab = "Number of Calls per Day", main = "Simulated Titi Calls",
     xlim = c(0, 30))
```

## How does your histogram compare to the shape of the probability mass function you plotted above?

The two plots look quite different! The probability mass function looks more like a perfect bell curve whereas my version of the simulated histogram almost shows to have a bimodal distribution (2 peaks). But this may be because I don't have enough data points. I think that if I have more than 104 samples, the data could look more like a normal distribution. We also need to keep in mind that these two graphs are showing very different things. The probability mass function graph plots the probabilities, not the individual data points (\# of calls).

> BP: Good, clear writeup of the differences between your histogram and the probability distribution. Mine ended up looking very similar though I often ended up with a much more obvious tail on the upper end of the distribution.

# Let's answer the probability questions now!

## What is the probability that she will hear more than 8 calls during any given session?

```{r}
# ppois get equal or less than -> so we need to get the complement probability
# This means 1 - P(x =< 8)
more_than_8_calls <- 1 - ppois(8, l)
more_than_8_calls
```

> BP: Looks good to me! I know you've already mentioned this in your peer commentary on my homework, but you can also set lower.tail = FALSE in the ppois function to get the upper tail (i.e. P(X \> x)), which saves you from having to subtract the result from 1.

## What is the probability that she will hear no calls in a session?

```{r}
no_calls <- dpois(0, l)
no_calls
```

## What is the probability that she will hear exactly 3 calls in a session?

```{r}
three_calls <- dpois(3, l)
three_calls
```

# Issues Encountered

1.  I initially misunderstood what the sampling simulation meant. I the output of a `dpois()` function were not data points but rather probabilities. However I decided to keep my mistake in the assignment answer just to remind me in the future
2.  Understanding what the differences between `dpois()` , `ppois()` , and `rpois()`
3.  Completed the initial assignment on Quarto but needed to convert to RMarkdown to apply the `cayman` package

> BP: Great job! Understanding the differences between dpois, ppois, and rpois was probably the biggest issue I ran into as well. Just a reminder too, but the HW instructions also say to find your favorite titi monkey picture and add it to the final document. I just browsed recent sightings in iNaturalist to find one that I liked!
