---
title: "AN588 - Homework 2"
author: "Christian Mei"
editor: visual
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: true          # Add a table of contents
    toc_depth: 2
---

# Titis in the Mist...

Every Saturday morning, at the same time, a primatologist goes and sits in the forest to listen for titi monkey calls, counting the number of calls they hear in a [2 hour window from 5am to 7am]{.underline}. Based on previous knowledge, she believes that the [mean number of calls she will hear in that time is exactly 15.]{.underline} Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

What is the probability that she will hear more than 8 calls during any given session?

What is the probability that she will hear no calls in a session?

What is the probability that she will hear exactly 3 calls in a session?

Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.

Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).

Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

# Answer Attempt

I decided to answer the questions in the order I think felt more inuitive

## **Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30**

We first need to create the distribution/dataset that the primatologist collected before plotting it.

```{r}
call_ranges <- 0:30
l = 15 # Poisson mean lambda based on mean number of calls
titis_calls <- dpois(x = call_ranges, lambda = l)


barplot(titis_calls, names.arg = call_ranges, space = 0, xlab = "Number of Calls per Day", ylab = "Pr(X = x)", main = "Probability Mass Function - Titis Calls")
```

## **Simulate 104 results from this distribution:**

My initial attempt which resulted in a mistake. Why do you think this is the case? Why can't I sample from the original `titis-calls` collection?

```{r}
sim <- NULL  # sets up a dummy variable to hold our 10000 simulations
n <- 104
for (i in 1:length(titis_calls)) {
    sim[i] <- mean(sample(titis_calls, n, replace = TRUE))
}

barplot(sim, names.arg = call_ranges, space = 0, xlab = "Number of Calls per Day", ylab = "Pr(X = x)", main = "Probability Mass Function - Titis Calls")


```

The problem above is that I am by sampling from `titis_calls` , I was sampling probability values and not data points (calls per day). Hence, this method is incorrect

What the assignment was actually asking for was to sample from a Poisson distribution with similar values, e.g same lambda mean = 15.

```{r}
simulated_calls <- rpois(104, lambda = l)

hist(simulated_calls, breaks = 0:30, freq = TRUE, col = "lightblue",
     xlab = "Number of Calls per Day", main = "Simulated Titi Calls",
     xlim = c(0, 30))
```

# **Let's answer the probability questions now!**

## What is the probability that she will hear more than 8 calls during any given session?

```{r}
# ppois get equal or less than -> so we need to get the complement probability
# This means 1 - P(x =< 8)
more_than_8_calls <- 1 - ppois(8, l)
more_than_8_calls
```

## What is the probability that she will hear no calls in a session?

```{r}
no_calls <- dpois(0, l)
no_calls
```

## What is the probability that she will hear exactly 3 calls in a session?

```{r}
three_calls <- dpois(3, l)
three_calls
```

# Issues Encountered

1.  I initially misunderstood what the sampling simulation meant. I the output of a `dpois()` function were not data points but rather probabilities. However I decided to keep my mistake in the assignment answer just to remind me in the future
2.  Understanding what the differences between `dpois()` , `ppois()` , and `rpois()`
3.  Completed the initial assignment on Quarto but needed to convert to RMarkdown to apply the `cayman` package
